\section*{Data and Code Availability}

All datasets analyzed in this study are public molecular-property datasets. The preprocessing, label-blind split generation, model training, calibration, conformal prediction, applicability-domain analysis, selective-prediction analysis, integrity audits, frozen result tables, manuscript figures, and LaTeX source files are maintained in the reproducibility repository: \url{https://github.com/DCarchimonde/aidd-paper-factory}. A submission-specific archival release and persistent identifier will be added before final submission.

\section*{Competing Interests}

The authors declare no competing interests.

\section*{Author Contributions}

S.T. conceptualized the study, designed and implemented the confirmatory benchmark, performed the analyses, audited the frozen outputs, prepared the figures, and drafted the manuscript. Y.W. reviewed the methodology, experimental design, interpretation, and manuscript consistency, and contributed to revision. Both authors approved the final manuscript.

\section*{AI-Assisted Technologies Statement}

During the preparation of this work, the authors used an AI-assisted tool to support language editing, manuscript organization, LaTeX formatting, code review, and reproducibility-material preparation. The authors reviewed and edited all content, verified the numerical results, code, figures, and references, and take full responsibility for the submitted manuscript.

\section*{Funding}

The authors received no specific funding for this work.

\section*{Acknowledgements}

The authors thank the maintainers of the public molecular datasets and the open-source cheminformatics, machine-learning, statistical, and scientific-computing libraries used in this work.
