\begin{abstract}
Molecular-property benchmarks are commonly summarized by aggregate predictive accuracy, yet such summaries may conceal chemically structured distribution shift, class-conditional failure, and uninformative uncertainty estimates. We evaluated reliability across four ADMET endpoints spanning binary classification (BBBP and ClinTox) and regression (ESOL and Lipophilicity). Frozen ECFP-based model families were assessed under label-blind random, scaffold, and similarity-cluster splits using repeated confirmatory seeds, explicit train--calibration--test separation, probability calibration, applicability-domain diagnostics, marginal and class-conditional conformal prediction, covariate-shift weighting, locally adaptive regression intervals, and retention-aware selective prediction. Random splitting generally produced the strongest predictive performance and probability calibration, while shift effects were endpoint dependent: ESOL performance deteriorated sharply under scaffold splitting, and mean Lipophilicity $R^2$ was negative under all three split designs. Chemical similarity was a stable risk and miscoverage indicator for BBBP and especially Lipophilicity, but was weak for ClinTox and reversed under the ESOL scaffold split. At nominal 90\% coverage, marginal and shift-weighted classification conformal methods concealed severe class-conditional failure; ClinTox positive-class coverage remained only 0.068--0.158 despite overall coverage near 0.90. Mondrian calibration restored positive-class coverage to approximately 0.93--0.94, but produced ambiguous two-label sets for approximately 71\% of samples. Shift-weighted and adaptive regression intervals showed no consistent universal coverage--efficiency advantage over standard marginal conformal prediction. Confidence-based rejection reduced ordinary ClinTox error partly by disproportionately removing positive examples, whereas similarity-based rejection preserved more minority cases but remained split dependent. These findings show that ADMET reliability is multidimensional: point performance, calibration, class-conditional validity, efficiency, chemical-domain diagnostics, and class-retention-aware selective risk must be interpreted jointly.
\end{abstract}

\noindent\textbf{Keywords:} ADMET prediction; applicability domain; conformal prediction; chemical distribution shift; calibration; selective prediction; class imbalance; chemometrics
