\section{Discussion}
\label{sec:discussion}

This study was designed to test whether common reliability summaries remain mutually consistent when molecular predictions are evaluated under chemical shift and class imbalance. They did not. Across four endpoints, point performance, probability calibration, chemical-domain distance, conformal coverage, prediction-set informativeness, and selective-risk ranking frequently pointed in different directions. The central implication is therefore methodological: molecular reliability is not a latent scalar that can be recovered by choosing a single preferred uncertainty score. It is a collection of conditional questions whose answers depend on the endpoint, shift mechanism, class composition, and intended use.

\subsection{Split design changed both performance and the meaning of uncertainty}

Random splitting produced the strongest average performance and calibration in this benchmark, but the difference from chemically structured splits was endpoint dependent. BBBP remained reasonably discriminative under scaffold and cluster evaluation, whereas ESOL deteriorated sharply under scaffold separation. This pattern agrees with recent molecular UQ studies showing that uncertainty rankings and apparent model reliability can change with the data-partitioning scenario \citep{koetter2025upgrading,tang2025polymeruq,parrondo2026shifts}. It also reinforces the rationale for benchmark protocols that compare methods on repeated, matched partitions and distinguish practical effect sizes from architecture-level averages \citep{ash2025comparison}.

The Lipophilicity results provide a particularly important boundary condition. Mean $R^2$ was negative under all three split designs, yet marginal conformal coverage remained close to 0.90. This is not a contradiction: split conformal calibration can construct wide enough intervals to attain marginal validity even when the point predictor has little explanatory value. It does, however, show why coverage cannot be used as a surrogate for predictive usefulness. Recent conformalized ADMET work has appropriately evaluated interval quality together with point error and width \citep{li2024conformalized}; our results extend that principle by showing that nominal validity can coexist with an unfavorable point-prediction baseline across every evaluated shift.

\subsection{Chemical similarity was informative, but not a universal uncertainty proxy}

Maximum training-set Tanimoto similarity was a reproducible risk and miscoverage indicator for Lipophilicity and a moderate indicator for BBBP. This finding is consistent with work showing that domain distance can identify difficult molecules when the representation and structure--property geometry are aligned with prediction error \citep{kim2025distance}. It also agrees with the broader observation that insufficient representation of a test molecule in training space contributes to molecular prediction error \citep{koetter2025upgrading}.

The counterexamples are at least as informative. Similarity was weak for ClinTox and reversed for ESOL under scaffold splitting. A scaffold-held-out molecule can be structurally novel according to one representation yet remain easy to predict because the target property depends on features shared across scaffolds; conversely, a close analogue can be difficult in a locally steep or noisy structure--property region. Recent molecular UQ benchmarks similarly report that no uncertainty measure dominates across datasets and that activity cliffs or steep local relationships can undermine distance-based rankings \citep{lanini2024unique,koetter2025upgrading,parrondo2026shifts}. Our threshold analysis further showed that a numerically convenient Tanimoto cutoff can absorb most of an endpoint into the ``low-domain'' group, leaving a small reference group and changing the contrast being estimated. Applicability-domain thresholds should therefore be treated as representation-specific operating rules, not universal chemical constants.

\subsection{Marginal coverage concealed a safety-relevant failure mode}

The most severe discrepancy occurred for ClinTox. Marginal and shift-weighted conformal prediction achieved overall coverage near the nominal 0.90 level while positive-class coverage remained between 0.068 and 0.158. This is a concrete instance of the distinction between marginal and conditional validity emphasized in conformal theory and subgroup-aware methods \citep{cauchois2021knowing,bastani2022multivalid,ding2023classconditional}. It also substantiates concerns raised in critical evaluations of binary conformal prediction: an aggregate coverage target can be met while errors are distributed very unevenly between classes \citep{krstajic2021critical}.

The practical relevance follows from the endpoint rather than from conformal theory alone. ClinTox positives represented only 6.45\% of the cleaned dataset, yet these are the observations associated with clinical toxicity. A procedure that covers the dominant negative class while almost systematically excluding the positive label is not rendered useful by a nominal marginal average. The appropriate minimum report for an imbalanced molecular classifier therefore includes positive- and negative-class coverage, class support, empty-set rate, and set size. Toxicity-focused conformal applications have demonstrated the promise of calibrated prediction sets \citep{zhang2021toxicity}; our results show why their evaluation must explicitly expose minority-class behavior.

\subsection{Mondrian calibration repaired validity by spending informativeness}

Mondrian calibration corrected ClinTox positive coverage across every frozen model--regime combination, reducing the class gap by approximately 0.77--0.83. That consistency is strong evidence that the marginal failure was not confined to one architecture. Yet the correction produced ambiguous two-label sets for roughly 71\% of test samples. In a binary task, a set containing both labels satisfies coverage while providing minimal discrimination. The result therefore illustrates the general coverage--precision trade-off in conformal prediction \citep{fisch2022trading}: statistical validity and operational informativeness are separate objectives.

This does not make Mondrian prediction a failed method. It identifies the decision it is making. When the minority calibration sample is small, achieving class-conditional coverage can require conservative thresholds. Such a set may still be valuable as an abstention or escalation signal, but it cannot be described as a precise toxicity decision. A deployment-oriented evaluation would need explicit downstream costs for false negatives, false positives, dual-label sets, and experimental follow-up. The present benchmark deliberately stops before assigning those application-specific utilities.

\subsection{Estimated shift correction and local adaptivity were not automatically beneficial}

Density-ratio weighting produced no stable class-level repair for ClinTox and no universal regression coverage--efficiency gain. This outcome should be interpreted in light of the method's assumptions. Weighted conformal prediction has a clear validity argument under covariate shift when the likelihood ratio is known, or estimated accurately enough for the target setting \citep{tibshirani2019covariate}. In molecular data, a classifier trained to distinguish calibration from test ECFP vectors may have limited effective sample size, may emphasize representation-specific differences, and cannot correct conditional label shifts or class imbalance merely by reweighting covariates. CoDrug demonstrates that density-aware molecular conformal prediction can improve coverage under designed distribution drifts \citep{laghuvarapu2023codrug}; our negative result shows that this should not be generalized to every endpoint, split, or ratio estimator.

The adaptive normalized regression baseline likewise produced only modest interval heterogeneity and weak width--error association. This matters because variable interval width is useful only when it tracks realized difficulty. Recent UQ benchmarks advocate evaluating ranking quality and calibration rather than assuming that architectural complexity produces better uncertainty \citep{lanini2024unique,jiang2024gnas,tang2025polymeruq}. Our paired results support that position: a more elaborate interval can be slightly narrower or more variable without delivering a reproducible joint improvement in coverage and efficiency.

\subsection{Selective prediction can improve a metric by changing who remains}

Confidence-based rejection reduced ordinary ClinTox risk but worsened balanced-risk pAURC and retained only about 27--31\% of positives when half of all samples were retained. The risk curve therefore combined two effects: ranking prediction difficulty and changing class composition. Similarity-based rejection retained more positives, but its balanced-risk benefit remained split dependent. This distinction parallels the motivation for validated, feature-aware, and coverage--precision-aware predictive inference \citep{cauchois2021knowing,fisch2022trading,teng2023feature}: a selective system must be evaluated on the composition and error structure of the accepted subset, not only its aggregate error.

For prospective screening, this result changes the question from ``Does abstention reduce error?'' to ``Which compounds and classes are removed to obtain that reduction?'' A confidence threshold that eliminates most toxicity positives may be unacceptable even if its ordinary accuracy improves. At minimum, imbalanced molecular selective-prediction studies should report the full retained-coverage curve, a matched random-rejection reference, balanced risk, and class-specific retention. A single operating point or aggregate AURC can conceal the mechanism of improvement.

\subsection{Implications for chemometric benchmark design}

The combined evidence supports a multidimensional reporting standard. First, random splits should be accompanied by at least one label-blind chemically structured evaluation. Second, repeated seeds should be summarized within model, and method contrasts should be paired on the same data partitions. Third, classification reports should separate discrimination, proper scoring rules, positive-probability calibration, and confidence calibration. Fourth, conformal validity should be reported both marginally and conditionally, together with set size, empty sets, and ambiguity. Fifth, applicability-domain claims should include continuous associations and threshold sensitivity. Sixth, selective prediction should report who remains, not only how low the retained risk becomes.

This framework complements rather than replaces recent molecular UQ toolkits and ADMET systems \citep{xiong2021admetlab,wei2022interpretable,swanson2024admetai,lanini2024unique}. Those resources broaden access to prediction and uncertainty estimation. The present contribution is an auditing layer for interpreting their outputs under repeated chemical shift and imbalance. Its main value lies in exposing conflicts that would disappear in a single leaderboard column.

\subsection{Limitations}

The study is limited to four public endpoints, ECFP4 representations, and four conventional model families. The observed failure modes cannot establish how message-passing networks, geometry-aware encoders, quantum-informed pretraining, or large molecular foundation models would behave \citep{heid2024chemprop,fang2022geometry,kim2024quantum}. At the same time, using a fixed representation reduced the risk that reliability conclusions were driven by changing the chemical feature space between diagnostics.

Confirmatory seed counts were finite, and several endpoint--domain groups were sparse. In particular, high-domain ClinTox groups under scaffold and cluster splitting could contain too few positive observations for stable class-specific metrics; those rows were retained as descriptive warnings rather than inferential evidence. The endpoint cleaning policy resolved duplicate classification labels by majority and removed exact ties, which is reproducible but cannot eliminate assay heterogeneity or annotation uncertainty. No wet-laboratory, temporal, institutional, or prospective external validation was performed. Finally, the weighted and adaptive conformal procedures represent specific prespecified implementations; failure to show a universal advantage does not establish that all shift-aware or locally adaptive methods will fail.

\subsection{Conclusion}

Trustworthy ADMET evaluation requires more than high aggregate accuracy or nominal marginal coverage. Chemical distribution shift, class imbalance, prediction-set informativeness, applicability-domain behavior, and selective-retention effects can point in different directions. The most defensible interpretation is therefore joint: point performance determines whether the underlying predictor is useful; calibration and conformal analysis assess different forms of probabilistic validity; efficiency determines whether valid outputs remain informative; applicability-domain diagnostics test whether chemical novelty ranks risk; and retention-aware selective analysis reveals which observations are removed to achieve lower error. Reporting these dimensions together provides a stronger chemometric basis for deciding when a molecular prediction should support, defer, or trigger downstream experimental work.
