\section{Introduction}
\label{sec:introduction}

Machine-learning models for absorption, distribution, metabolism, excretion, and toxicity (ADMET) prediction are commonly compared using aggregate discrimination or regression metrics. Such summaries are useful but incomplete. Molecular benchmarks are vulnerable to chemical distribution shift, model confidence can be poorly aligned with realized error, and class imbalance can make apparently reassuring averages insensitive to the compounds of greatest practical concern. These issues are especially important when models are evaluated on random splits that allow structurally similar compounds to appear in both training and test sets.

Probability calibration, conformal prediction, applicability-domain analysis, and selective prediction each address a different aspect of reliability. Calibration assesses the relationship between predicted probabilities and observed frequencies. Conformal prediction targets finite-sample marginal coverage under exchangeability assumptions. Applicability-domain diagnostics characterize the chemical relation between a query compound and the training set. Selective prediction evaluates whether uncertain cases can be rejected to reduce risk. None of these tools, considered alone, guarantees useful predictions under chemical shift or class imbalance.

A central difficulty is that aggregate validity can conceal subgroup failure. Marginal conformal coverage may remain close to its nominal target even when one class is severely under-covered. A method that restores class-conditional coverage may do so by returning broad or ambiguous prediction sets. Similarly, confidence-based rejection may reduce ordinary error while disproportionately excluding the minority class. These trade-offs require joint evaluation of validity, efficiency, chemical domain, and retained-sample composition.

This study develops a confirmatory chemometric evaluation of reliability across four ADMET endpoints spanning binary classification and regression. Frozen molecular model families were evaluated under label-blind random, scaffold, and similarity-cluster splits with explicit train--calibration--test separation and repeated confirmatory seeds. The analysis integrates predictive performance, probability calibration, continuous and threshold-based applicability-domain diagnostics, marginal and Mondrian conformal prediction, covariate-shift weighting, locally adaptive regression intervals, and retention-aware selective prediction.

The study addresses four research questions: (RQ1) how predictive performance and calibration change across random, scaffold, and cluster splits; (RQ2) whether chemical-domain distance identifies high-error or high-miscoverage compounds; (RQ3) whether marginal summaries conceal class-conditional or subgroup failure; and (RQ4) whether Mondrian, shift-weighted, or adaptive conformal methods improve practical reliability after accounting for efficiency and informativeness. The intended contribution is not a universally superior ADMET predictor, but a reproducible diagnostic framework showing when familiar reliability summaries agree, when they conflict, and what those conflicts imply for benchmark interpretation.

% Literature citations and the final journal-specific positioning paragraph will be
% inserted after the reference audit for the selected target journal.
