\section{Background and Related Work}

\subsection{Chemometric validation and sample partitioning}

Chemometric validation asks whether a model can generalize to observations that were not used, directly or indirectly, during model construction. The validity of that assessment depends on the relation between calibration and test samples, the intended prediction population, and the transparency of the validation design. Proper-validation principles distinguish resampling-based internal assessment from an independently meaningful test population \citep{esbensen2010proper}. Recent studies show that cross-validation and bootstrap can provide different views of robustness, that leakage can enter through sample selection or parameter tuning, and that validation strategies are often incompletely reported in applied chemometric studies \citep{lasfar2024robustness,kiraly2025leakage,ezenarro2025validation,camacho2026validation}. Sample partitioning is therefore not a clerical preprocessing step; it defines the chemical and statistical conditions under which performance is measured.

Applicability-domain research provides a complementary perspective. Predictions tend to be more reliable when test compounds remain within the chemical or descriptor space represented by the training data \citep{netzeva2005applicability,sahigara2012applicability}. A test set that is too close to the training set can overstate extrapolative performance, whereas a test set that is distant in several uncontrolled ways can confound chemical novelty with response shift or sparse coverage. Validation benefits from reporting these dimensions separately.

\subsection{Benchmarks, activity cliffs, and split-induced difficulty}

MoleculeNet, Therapeutics Data Commons, and OGB-LSC have made molecular machine-learning studies more comparable by organizing public datasets, tasks, metrics, and common evaluation settings \citep{wu2018moleculenet,huang2021tdc,hu2021ogblsc}. Benchmark extensions have also emphasized class imbalance and activity-cliff behaviour, because average scores can hide failures on low-data, imbalanced, or locally discontinuous regions of chemical space \citep{li2022imdrug,jimenez2022attribution,vantilborg2022cliffs,zhang2023acnet}.

Random splitting is statistically convenient and often preserves the overall target distribution, but it can place near-neighbour compounds in both training and test sets. The resulting score is informative for interpolation-like prediction, yet it may not represent transfer to a new chemical series. Scaffold splitting is widely used as a stricter alternative. Molecules are grouped by frameworks related to Bemis--Murcko scaffolds, and complete groups are assigned to one side of the partition \citep{bemis1996properties}. This prevents direct scaffold overlap, but it does not guarantee chemical dissimilarity across the partition. Recent analyses have shown both that learned molecular representations retain important limitations on small benchmark datasets and that nominally different scaffolds may remain chemically similar across a scaffold split \citep{dias2023limitations,guo2024scaffold}. Holding out complete scaffold groups can also create shifted target distributions or small, concentrated test sets.

\subsection{Contemporary molecular representation learning}

Recent molecular-property models span one-dimensional chemical language models, two-dimensional graph encoders, three-dimensional geometric models, and multimodal combinations. SMILES-based and language-model approaches include multi-granularity sequence learning, supervised grammar-based latent models, and large-scale chemical transformers \citep{jiang2022multigran,oliveira2022sgvae,ross2022molformer,sultan2024transformers}. Two-dimensional graph representation learning has progressed through contrastive pretraining, knowledge-guided graph transformers, hierarchical self-supervision, chemical-substructure explanations, knowledge-enhanced contrastive learning, molecule-level graph structure learning, and knowledge-enhanced transformers \citep{wang2022molclr,li2023kpgt,zang2023hierarchical,wu2023substructure,fang2023kgcontrastive,zhao2024gslmpp,gao2024knomol}.

Three-dimensional and multimodal approaches incorporate conformational geometry, cross-view pretraining, image-derived representations, pharmacophoric constraints, structure--text alignment, and joint 1D/2D/3D learning \citep{fang2022gem,li2022adaptive,zeng2022image,liu2022graphmvp,zhu2022unified,jiang2023pharmacophore,zhou2023unimol,liu2023moleculestm,luo2025molclsp}. These studies demonstrate rapid architectural progress, but they also rely on benchmark partitions whose chemical proximity, target distributions, and scaffold composition may differ. Consequently, a higher score can reflect both model capability and the validation condition under which the score was produced.

\subsection{Diagnostic rather than leaderboard evaluation}

Large molecular representation studies and recent reviews show that performance rankings vary across endpoints, model families, data regimes, and evaluation choices \citep{yang2019learned,feinberg2018potentialnet,vantilborg2022cliffs,dias2023limitations,sultan2024transformers}. This motivates an evaluation design that reports how the partition changes the test problem before attributing a score difference to the model.

The present study follows this diagnostic view. Random and scaffold protocols are not treated as points on a universal scale of difficulty. Instead, each partition is characterized by target shift, scaffold composition, chemical similarity, predictive gap, and ranking behaviour. A target-balanced scaffold partition is added as a controlled counterfactual to isolate the contribution of first-moment target shift while retaining scaffold disjointness.